k-means clustering with penguins

### 2020-07-09 code_tests/kmeans_testing.Rmd
add k-means clustering & text edits (#7)
* actually deal with conflicts

* update draft w/kmeans

* push recent code ?

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(shadowtext)
library(NbClust)
library(patchwork)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

@@

The "old" way, using `kmeans()`:
```{r}
penguins_k <- penguins %>% 
  filter(sex == "male") %>% 
  select(species, bill_length_mm:body_mass_g) %>% 
  drop_na()

# Using bill_length & flipper_length
# THIS ONE SHOWS *pretty* good clustering

bills <- penguins_k %>% select(species, bill_length_mm, bill_depth_mm)

set.seed(1967)
kclust_bills <- kmeans(bills[,2:3], centers = 3, nstart = 20)
kclust_bills

# Table of assigned clusters for penguins
bills_class <- table(kclust_bills$cluster,bills$species)%>% 
  as.data.frame() %>% 
  rename(cluster = Var1, species = Var2, n = Freq) %>% 
  pivot_wider(names_from = species, values_from = n)

# Same for iris:

petals <- iris %>% select(Species, Petal.Length, Petal.Width)

set.seed(1967)
kclust_petals <- kmeans(petals[,2:3], centers = 3, nstart = 25)
kclust_petals

# Table of assigned clusters for iris petals: 
petal_class <- table(kclust_petals$cluster, petals$Species) %>% 
  as.data.frame() %>% 
  rename(cluster = Var1, Species = Var2, n = Freq) %>% 
  pivot_wider(names_from = Species, values_from = n)

```

Visualizing it: 
```{r}
bills_cluster <- cbind(bills, cluster_no = kclust_bills$cluster)

penguins_kmeans <- ggplot(data = bills_cluster, aes(x = bill_length_mm, 
                                     bill_depth_mm)) +
  geom_text(aes(label=cluster_no, color = species), 
                  size = 4) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  labs(title = "Penguins k-means clustering") +
  theme(legend.position = "bottom")

iris_cluster <-cbind(petals, cluster_no = kclust_petals$cluster)

iris_kmeans <- ggplot(data = iris_cluster, aes(x = Petal.Length, y = Petal.Width)) +
  geom_text(aes(label = cluster_no, color = Species), size = 4) +
  scale_color_manual(values = c("darkorange","darkorchid","cyan4")) +
  theme(panel.background = element_rect(fill = "gray90", color = "gray90"),
        legend.position = "bottom") +
  labs(title = "Iris k-means clustering")

(penguins_kmeans | iris_kmeans)

# Full table of cluster assignments: 
combo_cluster <- cbind(bills_class, petal_class) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)

combo_cluster

```
## Using tidymodels? eeeeehhh we'll see

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(shadowtext)
library(NbClust)
library(patchwork)
library(kableExtra)
library(mclust)
knitr::opts_chunk$set(echo = TRUE)
```

@@

Using `kmeans()`:

Example 1: Male penguins bill dimensions
```{r}
mp_bills <- penguins %>% 
  filter(sex == "male") %>% 
  select(species, bill_length_mm, bill_depth_mm)

mp_bills_nospecies <- mp_bills %>% 
  select(-species) %>% 
  scale() %>% 
  as.data.frame()

# Perform k-means on male penguin bill dimensions (k = 3, w/20 centroid starts)
mp_bills_kmeans <- kmeans(mp_bills_nospecies, centers = 3, nstart = 20)
mp_bills_kmeans

# Bind clusters to the data
mp_kmeans_clusters <- cbind(mp_bills_nospecies, cluster = mp_bills_kmeans$cluster, species = mp_bills$species)

ggplot(mp_kmeans_clusters, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_text(aes(label = cluster, color = species))

male_cluster_counts <- mp_kmeans_clusters %>% 
  complete() %>% 
  count(species, cluster) %>% 
  pivot_wider(names_from = species, values_from = n)

male_cluster_counts
# SERIOUSLY WTF...those 3s are NOT closer to the 3 centroid...
# 
# plot(mp_bills_nospecies, col = mp_bills_kmeans$cluster)
# points(mp_bills_kmeans$centers, col = 1:2, pch = 8, cex = 2)
# 
# # Testing some tidymodels methods instead...
# test_cluster <- augment(mp_bills_kmeans, mp_bills_nospecies) %>% 
#   rename(cluster_number = .cluster)
# 
# ggplot(data = test_cluster, aes(x = bill_length_mm, y = bill_depth_mm, color = cluster_number)) +
#   geom_point()
```

# With all penguins (bill dimensions)

Example 2: All penguins bill dimensions
```{r}
p_bills <- penguins %>% 
  select(species, bill_length_mm, bill_depth_mm) %>% 
  drop_na()

p_bills_nospecies <- p_bills %>% 
  select(-species) %>% 
  scale() %>% 
  as.data.frame()

# Perform k-means on male penguin bill dimensions (k = 3, w/20 centroid starts)
p_bills_kmeans <- kmeans(p_bills_nospecies, centers = 3, nstart = 20)
p_bills_kmeans

# Bind clusters to the data
p_kmeans_clusters <- cbind(p_bills_nospecies, cluster = p_bills_kmeans$cluster, species = p_bills$species)

ggplot(p_kmeans_clusters, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_text(aes(label = cluster, color = species))

p_cluster_counts <- p_kmeans_clusters %>% 
  complete() %>% 
  count(species, cluster) %>% 
  pivot_wider(names_from = species, values_from = n)

p_cluster_counts
```

# Trying EM using `mclust` package for fun

See: https://en.wikibooks.org/wiki/Data_Mining_Algorithms_In_R/Clustering/Expectation_Maximization_%28EM%29#Executing_the_Algorithm

```{r}
BIC <- mclustBIC(mp_bills_nospecies)
plot(BIC)
summary(BIC)

mod1 <- Mclust(mp_bills_nospecies, G = 3, x = BIC)
summary(mod1, parameters = TRUE)

plot(mod1, what = "classification")
```
